Paper template for {ICPhS} 2023 Prague

### 2024-04-23 manuscript/main.Rmd

Both Catalan and Peninsular Spanish sound inventories have an phoneme /s/, which can be classified as a voiceless alveolar fricative (Hualde & Colina, 2015). In addition, most of the Catalan varieties have a voiced alveolar phoneme in intervocalic contexts (Lloret, 2011; Recasens Vives, 1996, p. 279). In onset position, /s/ and /z/ contrast phonemically, which leads to minimal pairs like *cel* [s\textepsilon l] - *zel* [z\textepsilon l] and *fossa* [fósa] - *fosa* [fóza] (Carrera-Sabaté & Prieto, 2012-2019). Conversely, at the end of the word, this contrast is neutralized, being the voicing or devoicing of the alveolar fricative dependent on the following sound (Institut d’Estudis Catalans, 2022). For example, if it is followed by a vowel, the alveolar fricative is voiced: *torres* [tór\textschwa{}s] and *torres antigues* [tor\textschwa{}z\textschwa{}nti\textgamma{}\textschwa{}s].

Interestingly, Medieval Spanish, similarly to Modern Catalan, distinguished between /s/ and /z/ in intervocalic position within a word, although these sounds merged through a devoicing process (Hualde & Colina, 2015). Nevertheless, in those modern varieties where the /s/ in coda position is produced, such as in Peninsular Spanish, this voiceless phoneme might assimilate to the following sound, especially if it is a voiced consonant (*mismo*, [mízmo]), rather than a vowel (*las alas*, [lasálas]) (Hualde & Colina, 2015; Hualde et al., 2020). Hence, Spanish inventory does not have the phoneme /z/, but [z] may occur as an allophone of /s/, due to a voice assimilation process. It is important to mention, though, that it has been claimed that this voicing is generally not complete, but partial (Benet et al., 2012; Hualde & Colina, 2015; Hualde et al., 2020).

There is some evidence of the influence of Spanish on the production of the Catalan alveolar fricative, in contexts where, according to prescriptivist rules (Institut d'Estudis Catalans, 2022), the [z] should be produced, but [s] is produced instead. This phenomenon has mainly been attested in those areas where there has been an extended contact between Catalan and Spanish, such asthe so-called "franja de Aragón" (borderland between Catalonia and Aragon), some areas of the Valencian Community, and Barcelona and other urban areas (Benet et al., 2012; Davidson, 2020; Recasens Vives, 1996, p. 268). The opposite pattern has also been observed - that is, the voicing of the alveolar fricative in the Spanish spoken in Barcelona and Valencia. Interestingly, although in both Barcelona and Valencia Catalan and Spanish coexist and influence each other, it seems that Catalan has a greater influence on Spanish in Barcelona, but  Spanish has a greater influence on Catalan in Valencia (Davidson, 2020).

@@

The social meaning and the perception of intervocalic [z] in Spanish has been studied in a variety of contexts, including both monolingual (Chappell, 2016, 2017, 2021; Chappell et al., 2023; Chappell & García, 2017; García, 2015, 2019; Silverstein, 2003) and bilingual contexts (Davidson, 2014, 2015, 2019; Sinner, 2002), despite having been categorized as a low-salient variant (Chappell, 2017, 2021; Chappell et al., 2023; García, 2015)With respect to the former, it has been found that for listeners from San José, Costa Rica, the covert attitudes associated with intervocalic [z] vary significantly by speaker gender. Consequently, male speakers who use [z] tend to be positively evaluated, and generally perceived as nice, local, confident, and masculine. Conversely, female speakers who use [z] are only afforded negative associations of lower education and social class (Chappell, 2016; Chappell et al., 2023; Silverstein, 2003).

Similarly, García (2019) shows that, in Ecuador, intervocalic /s/ is evaluated differently in male and female speakers. Accordingly, women who produce intervocalic [z] are viewed as significantly lower status, less pleasant, and older. Contrary, men who produce intervocalic [z] do not receive these negative evaluations. Additionally, this variable has been found to serve as a marker of regional identity, with voicing being associated with neighboring Highland Ecuadorian dialects (Cuenca and Quito) that exhibit systematic voicing. García (2020) hypothesizes that young men are the leaders of this change, essentially due to the fact that they are in greater contact with these highland varieties. Interestingly, the mean voicing rates of females are roughly two age groups behind those of males. Consequently, it is believed that it is not necessarily the case that women in Ecuador are avoiding voicing like San José women due to its negative evaluations, but rather that they are further behind in the adoption of the incoming variant.

With respect to the social meaning and the perception of intervocalic [z] in bilingual contexts, this phenomenon has been mainly studied in Catalan-Spanish contact situations, given its robust usage by Catalan-Spanish bilinguals and, in particular, by those with greatest exposure to and usage of Catalan (Davidson, 2014, 2015). According to Sinner (2002), intervocalic [z] can be considered a possible linguistic marker of Catalonian Spanish, despite its relative lack of social awareness (Davidson, 2019; Sinner, 2002). In their study, Sinner (2002) interviewed 12 speakers of Barcelonan Spanish and monolingual Spanish-speakers from Madrid aged 27 to 41 regarding their awareness of linguistic features of Catalanized Spanish. Results showed that only select Barcelona speakers perceived intervocalic [z] as a linguistic marker of Catalonian Spanish, whereas monolingual Spanish-speakers from Madrid did not. Additionally, Davidson (2019) found that, in Barcelona, the indexical field of intervocalic [z] features exclusively positive associations with bilingualism and local solidarity, which warrants its gradual adoption in the speech community, unlike other phonetic markers of Catalonian Spanish that are the topic of overt social ridicule, such as lateral velarization. These results are in line with previous findings on intervocalic [z] in Catalonian Spanish. For instance, Silverstein (2003) found that intervocalic [z] indexes a local Barcelona identity, and the variant then acquires positive associations with bilingualism and cultural pride. Overall, these findings suggest that intervocalic [z] likely functions as a sociolinguistic indicator of Catalonian Spanish within the Catalonian speech community, indexing Catalan-Spanish bilingual membership while lacking any considerable degree of generalized conscious awareness (Davidson, 2019).

>>>>

Both Catalan and Peninsular Spanish sound inventories have an phoneme /s/, which can be classified as a voiceless alveolar fricative \cite{hualde_colina_2015}. In addition, most of the Catalan varieties have a voiced alveolar phoneme in intervocalic contexts **(Lloret, 2011; Recasens Vives, 1996, p. 279)**. In onset position, /s/ and /z/ contrast phonemically, which leads to minimal pairs like *cel* [s\textepsilon l] - *zel* [z\textepsilon l] and *fossa* [fósa] - *fosa* [fóza] **(Carrera-Sabaté & Prieto, 2012-2019)**. Conversely, at the end of the word, this contrast is neutralized, being the voicing or devoicing of the alveolar fricative dependent on the following sound **(Institut d’Estudis Catalans, 2022)**. For example, if it is followed by a vowel, the alveolar fricative is voiced: *torres* [tór\textschwa{}s] and *torres antigues* [tor\textschwa{}z\textschwa{}nti\textgamma{}\textschwa{}s].

Interestingly, Medieval Spanish, similarly to Modern Catalan, distinguished between /s/ and /z/ in intervocalic position within a word, although these sounds merged through a devoicing process \cite{hualde_colina_2015}. Nevertheless, in those modern varieties where the /s/ in coda position is produced, such as in Peninsular Spanish, this voiceless phoneme might assimilate to the following sound, especially if it is a voiced consonant (*mismo*, [mízmo]), rather than a vowel (*las alas*, [lasálas]) **(Hualde & Colina, 2015; Hualde et al., 2020)**. Hence, Spanish inventory does not have the phoneme /z/, but [z] may occur as an allophone of /s/, due to a voice assimilation process. It is important to mention, though, that it has been claimed that this voicing is generally not complete, but partial **(Benet et al., 2012; Hualde & Colina, 2015; Hualde et al., 2020)**.

There is some evidence of the influence of Spanish on the production of the Catalan alveolar fricative, in contexts where, according to prescriptivist rules **(Institut d'Estudis Catalans, 2022)**, the [z] should be produced, but [s] is produced instead. This phenomenon has mainly been attested in those areas where there has been an extended contact between Catalan and Spanish, such asthe so-called "franja de Aragón" (borderland between Catalonia and Aragon), some areas of the Valencian Community, and Barcelona and other urban areas **(Benet et al., 2012; Davidson, 2020; Recasens Vives, 1996, p. 268)**. The opposite pattern has also been observed - that is, the voicing of the alveolar fricative in the Spanish spoken in Barcelona and Valencia. Interestingly, although in both Barcelona and Valencia Catalan and Spanish coexist and influence each other, it seems that Catalan has a greater influence on Spanish in Barcelona, but  Spanish has a greater influence on Catalan in Valencia **(Davidson, 2020)**.

@@

The social meaning and the perception of intervocalic [z] in Spanish has been studied in a variety of contexts, including both monolingual **(Chappell, 2016, 2017, 2021; Chappell et al., 2023; Chappell & García, 2017; García, 2015, 2019; Silverstein, 2003)** and bilingual contexts **(Davidson, 2014, 2015, 2019; Sinner, 2002)**, despite having been categorized as a low-salient variant **(Chappell, 2017, 2021; Chappell et al., 2023; García, 2015)**. With respect to the former, it has been found that for listeners from San José, Costa Rica, the covert attitudes associated with intervocalic [z] vary significantly by speaker gender. Consequently, male speakers who use [z] tend to be positively evaluated, and generally perceived as nice, local, confident, and masculine. Conversely, female speakers who use [z] are only afforded negative associations of lower education and social class **(Chappell, 2016; Chappell et al., 2023; Silverstein, 2003)**.

Similarly, **García (2019)** shows that, in Ecuador, intervocalic /s/ is evaluated differently in male and female speakers. Accordingly, women who produce intervocalic [z] are viewed as significantly lower status, less pleasant, and older. Contrary, men who produce intervocalic [z] do not receive these negative evaluations. Additionally, this variable has been found to serve as a marker of regional identity, with voicing being associated with neighboring Highland Ecuadorian dialects (Cuenca and Quito) that exhibit systematic voicing. **García (2020)** hypothesizes that young men are the leaders of this change, essentially due to the fact that they are in greater contact with these highland varieties. Interestingly, the mean voicing rates of females are roughly two age groups behind those of males. Consequently, it is believed that it is not necessarily the case that women in Ecuador are avoiding voicing like San José women due to its negative evaluations, but rather that they are further behind in the adoption of the incoming variant.

With respect to the social meaning and the perception of intervocalic [z] in bilingual contexts, this phenomenon has been mainly studied in Catalan-Spanish contact situations, given its robust usage by Catalan-Spanish bilinguals and, in particular, by those with greatest exposure to and usage of Catalan **(Davidson, 2014, 2015)**. According to **Sinner (2002)**, intervocalic [z] can be considered a possible linguistic marker of Catalonian Spanish, despite its relative lack of social awareness **(Davidson, 2019; Sinner, 2002)**. In their study, Sinner (2002) interviewed 12 speakers of Barcelonan Spanish and monolingual Spanish-speakers from Madrid aged 27 to 41 regarding their awareness of linguistic features of Catalanized Spanish. Results showed that only select Barcelona speakers perceived intervocalic [z] as a linguistic marker of Catalonian Spanish, whereas monolingual Spanish-speakers from Madrid did not. Additionally, Davidson (2019) found that, in Barcelona, the indexical field of intervocalic [z] features exclusively positive associations with bilingualism and local solidarity, which warrants its gradual adoption in the speech community, unlike other phonetic markers of Catalonian Spanish that are the topic of overt social ridicule, such as lateral velarization. These results are in line with previous findings on intervocalic [z] in Catalonian Spanish. For instance, **Silverstein (2003)** found that intervocalic [z] indexes a local Barcelona identity, and the variant then acquires positive associations with bilingualism and cultural pride. Overall, these findings suggest that intervocalic [z] likely functions as a sociolinguistic indicator of Catalonian Spanish within the Catalonian speech community, indexing Catalan-Spanish bilingual membership while lacking any considerable degree of generalized conscious awareness **(Davidson, 2019)**.

### 2024-04-23 manuscript/main.tex

Both Catalan and Peninsular Spanish sound inventories have an phoneme
/s/, which can be classified as a voiceless alveolar fricative (Hualde
\& Colina, 2015). In addition, most of the Catalan varieties have a
voiced alveolar phoneme in intervocalic contexts (Lloret, 2011; Recasens
Vives, 1996, p.~279). In onset position, /s/ and /z/ contrast
phonemically, which leads to minimal pairs like \emph{cel}
{[}s\textepsilon l{]} - \emph{zel} {[}z\textepsilon l{]} and
\emph{fossa} {[}fósa{]} - \emph{fosa} {[}fóza{]} (Carrera-Sabaté \&
Prieto, 2012-2019). Conversely, at the end of the word, this contrast is
neutralized, being the voicing or devoicing of the alveolar fricative
dependent on the following sound (Institut d'Estudis Catalans, 2022).
For example, if it is followed by a vowel, the alveolar fricative is
voiced: \emph{torres} {[}tór\textschwa{}s{]} and \emph{torres antigues}
{[}tor\textschwa{}z\textschwa{}nti\textgamma{}\textschwa{}s{]}.

>>>>

Both Catalan and Peninsular Spanish sound inventories have an phoneme
/s/, which can be classified as a voiceless alveolar fricative
\cite{hualde_colina_2015}. In addition, most of the Catalan varieties
have a voiced alveolar phoneme in intervocalic contexts \textbf{(Lloret,
2011; Recasens Vives, 1996, p.~279)}. In onset position, /s/ and /z/
contrast phonemically, which leads to minimal pairs like \emph{cel}
{[}s\textepsilon l{]} - \emph{zel} {[}z\textepsilon l{]} and
\emph{fossa} {[}fósa{]} - \emph{fosa} {[}fóza{]} \textbf{(Carrera-Sabaté
\& Prieto, 2012-2019)}. Conversely, at the end of the word, this
contrast is neutralized, being the voicing or devoicing of the alveolar
fricative dependent on the following sound \textbf{(Institut d'Estudis
Catalans, 2022)}. For example, if it is followed by a vowel, the
alveolar fricative is voiced: \emph{torres} {[}tór\textschwa{}s{]} and
\emph{torres antigues}
{[}tor\textschwa{}z\textschwa{}nti\textgamma{}\textschwa{}s{]}.
